---
title: "EEB 5301 Homework 4"
author: "Samantha Apgar & Henry Frye"
date: "Spring 2017"
output: 
  tufte::tufte_handout:
    keep_tex: false
---

# Problem 1
 
The students' suggestion of doubling plant densities will not allow the aphid population to grow amidst ladybug predation. The predator prey relationship was unstable at a carrying capacity of 1,000 aphids and remains unstable at a carrying capacity of 1,200 aphids. The phase diagram below shows that the predator isocline remains to the left of both scenarios, indicating an unstable predator prey relationship. 

```{r}
r=.001
a=.5
k1=1000
k2=1200
h=.02083
q=1/3
f=.01

N1= 1:k1
N2= 1:k2
Prey1 = (r*N1 * (1 - (N1/k1))* (a*N1)) / (1+a*h*N1)
Prey2 = (r*N2 * (1 - (N2/k2))* (a*N2)) / (1+a*h*N2)
#Prey2 = (r / (a * k2)) * (k2 - (k2*a*h + 1)*N2 - a*h*(N2^2))

Predator = ( q / (f*a - a*h*q))

plot(x = N1, y  = Prey1, bty = 'n', ylim = c(0,15), type = 'l')
abline(v = Predator, col = 'red')
lines(Prey2, col = 'cadetblue')

```

# Problem 2

```{r vegan setup, echo = FALSE, warning = FALSE, message = FALSE}
library(vegan)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggthemes)
data(dune)
data(dune.env)

dune.sum	<- aggregate(dune,	list(Moisture	=	dune.env$Moisture),	sum)
dune.sum1	<- dune.sum[,-1]
rownames(dune.sum1)	<- dune.sum[,1]

radPlot	<- function(x,	add=FALSE,	col='black',	...){
		x	<- x[x>0]
		x	<- x[rev(order(x))]
		if(!add)
				plot(1:length(x),	x,	type='l',	xlab='Rank',	ylab='Abundance	(log	scale)',
									log='y',	col=col,	...)
		else
				lines(1:length(x),	x,	col=col,	...)
		return(NULL)}

```

##Part a

```{r}
specnumber(dune.sum1)
```

As moisture increases, species richness decreases from moisture level 1 to level 4, but then suddenly increases at level 5. The most rich is level 5 while level 4 is the least rich.

## Part b

```{r}
sort(diversity(dune.sum1))
```

According to Shannon's Index, the moisture level is ranked 5 > 2 > 1 > 4 in order of high diversity to lowest.

## Part c

```{r}
sort(exp(diversity(dune.sum1)))
sort(1/(diversity(dune.sum1, index = 'invsimpson')))
```

The effective number of species between Simpson's and Shannon's index are almost flipped.  Moisture level 4 has the lowest effective number of species according to Shannon's index while the level 4 is considered the having the highest effective number of species according to Simpson's.

## Part d

```{r}
diversity(dune.sum1[1,], index = 'invsimpson')
```

Taking the inverse of Simpson's index yields the probability of finding the same two species in a sample, in this case this 7.8%.

## Part e


```{r}
radPlot(x = dune.sum1[1,], bty = 'n')
radPlot(x = dune.sum1[2,], add = TRUE, col = 'blue')
radPlot(x = dune.sum1[3,], add = TRUE, col = 'red')
radPlot(x = dune.sum1[4,], add = TRUE, col = 'limegreen')
legend('topright', legend = c('1','2',"4","5"),
       col = c("black","blue","red","limegreen"), lty =1,lwd=2)
```


## Part f

Based on the plot above, moisture level 1, appears to be the most even since it has the flattest slope. However, this is is not much flatter than level 5 and not easily interpreted.

One way to better quantify evenness would be to calculate the Hill's ratio from either of the diversity indices above.

## Part g

Simpson's index tends to be less sensitive to rare species than Shannon's index. It is weighted towards the most abundant species.  In part c, Simpson's index tended to favor level 4 as the most diverse, which makes sense from the graph in part e that shows level 4 as the most dominated by common abundant species (i.e. lack of sensitivity towards rare species).

## Part h
```{r}
library(knitr)
rarefaction_data <- data.frame("Moisture Level 1" = c(rarefy(dune.sum1[1,], sample =2),
                                rarefy(dune.sum1[1,], sample =10),
                                rarefy(dune.sum1[1,], sample =100)),
           "Moisture Level 2" = c(rarefy(dune.sum1[2,], sample =2),
                                rarefy(dune.sum1[2,], sample =10),
                                rarefy(dune.sum1[2,], sample =100)),
           "Moisture Level 4" = c(rarefy(dune.sum1[3,], sample =2),
                                rarefy(dune.sum1[3,], sample =10),
                                rarefy(dune.sum1[3,], sample =100)),
           "Moisture Level 5" = c(rarefy(dune.sum1[4,], sample =2),
                                rarefy(dune.sum1[4,], sample =10),
                                rarefy(dune.sum1[4,], sample =100))
)

rownames(rarefaction_data) <- c("2","10","100")

kable(rarefaction_data)
```


The expected number of species for each moisture level are listed in the table above. Different sampling efforts within different moisture levels may yield more rare species thus affecting the overall diversity metrics (especially depending which index is used).


## Part i

```{r}
vegdist(dune.sum1)
```

Based on the dissimilarity index table above, the farther the distance between moisture, the higher difference in indices.  For example, when moisture level 1 is compared to the others the distance index increase with increases difference between moisture.
