---
title: "Population and Community Ecology Homework 2"
author: Samantha Apgar & Henry Frye
output: pdf_document

---

# Part 1

##Tutorial Run-Through

```{r Analysis of Point Patterns}
# install.packages("spatstat")
library(spatstat)

data(japanesepines)
summary(japanesepines)

plot(japanesepines)

clarkevans(japanesepines)

clarkevans.test(japanesepines, correction = "Donnelly")

RipleysK <- Kest(japanesepines)
plot(envelope(japanesepines))

RipleysL <- Lest(japanesepines)
plot(envelope(japanesepines, Lest, nsim = 99), .-theo~r)

p <- pcf(japanesepines)
plot(p)

plot(envelope(japanesepines, pcf, nsim = 99))
```

```{r Quadrats and the Poisson Distribution}

data(longleaf)
summary(longleaf)

QC <- quadratcount(longleaf)
QC
  
plot(longleaf)
plot(QC, add = TRUE, cex = 1.5, col = "red", font = 2)

QC <- quadratcount(longleaf, 15, 15)
QC

quadrat.test(longleaf, 15, 15, method = "M", nsim = 1999)

```

## Part 1a. 

```{r Problem Set: White Oaks}

whiteoaks <- lansing[lansing$marks == "whiteoak"]

clarkevans(whiteoaks)
clarkevans.test(whiteoaks, correction = "Donnelly")


```

The Clark Evans Test with Donnelly correction yields a p-value of 0.01 (p < .05). This indicates that we can reject the null hypothesis that the data is randomly distributed. 

## Part 1b. 

Yes, the Clark and Evans Index is significantly different from 1. Our R value is 0.93346 and an R value < 1 indicates spatial clumping. 

## Part 1c. 

```{r}
RipleysL <- Lest(whiteoaks)
plot(envelope(whiteoaks, Lest, nsim = 99), .-theo~r)

```

Because the observed Ripley's $\hat{L}$ is outside of the expected confidence envelope, we can conclude that the spatial distribution deviates form a normal distribution.

## Part 1d. 

```{r}
<<<<<<< HEAD
QC <- quadratcount(whiteoaks)
QC
  
=======
>>>>>>> dcf834a9d9797abb5cb3342297b27b468c7b4cc1
plot(whiteoaks)
plot(QC, add = TRUE, cex = 1.5, col = "red", font = 2)

QC <- quadratcount(whiteoaks, 10, 10)

quadrat.test(whiteoaks, 10, 10, method = "M", nsim = 1999)
```

## Part 1e. 
The p-value of the quadrat test is 0.001 (p < .05), indicating that we can reject the null hypothesis that the distribution of white oaks is random. The data indicates that the white oaks are not randomly distributed throughout the plot and may be either clumped or uniform. Visually, it looks as if the white oak distribution is clumped. 



