---
title: "Population & Community Ecology Homework 1"
author: Samantha Apgar & Henry Frye
output: 
  pdf_document: default
---

#Problem 1

In this problem we look to find the density of duckweed plants per $m^2$ after 47 days. We will estimate the future population using the exponential growth formula, $N_t = N_0 e^{rt}$ where the initial population, $N_0$, is 11.1 duckweed plants per $m^2$.  We can calculate the rate of unitless population change, r, by dividing the estimate after 20 day from the initial population.

```{r}
r = 23.5/11.1
r
```

Thus with r= 2.12, we can calculate $N_{27}$ as $N_{27} = 11.1 e^{2.12\times47}$.

```{r}
n27 <- 11.1*exp(2.12*47)
n27
```

Thus after a total of 47 days, the population density will be $2.08\times10^{44}$ duckweed plants per $m^2$.

***

#Problem 2



***

#Problem 3

##Part a 
Given the information we would like to solve for $N_7$, the population in seven years.

```{r}
n_7 <- (1.037^7)*112
n_7
```

Thus the population density will be 144 (rounded, no units given) at year 7.

##Part b
Using the exponential growth formula, $N_t = N_0 e^{rt}$, we would like to
solve for an r value where $N_0 = 112$ and $N_t = 144$. Thus we solve for r as $ln(\frac{N_t}{N_0})/t=r$.  

```{r}
r <- log(144/112)/7
r
```

Thus r is equal to .036 in the exponential model to  achieve the same results as the discrete population growth model.

***

#Problem 4

***



#Problem 5

Survivorship, as defined by $\frac{S_x}{S_0}$, is calculated in the vector below.
```{r}
age <- 0:4
sx <- c(740,280,105,32,0)
bx <- c(0,0.4,1.3,3,8,NA)
lx <- sx/740
print(lx)
```

**I assume that g(x) was meant to be p(x)**

Solving for p(x)


***

#Problem 6


***

#Problem 7

The following is the setup for the population project matrix:
```{r}
stages <- c('seedling','juvenile','small adult', 'medium adult', 'large adult')
A <-  matrix(c(
     0,  0,.015,.063,.189,
  .294,.25, 0  , 0  , 0  ,
  .176,.35,.385,.059,.012,
  0   ,.35,.461,.568,.171,
  0   , 0 ,.108,.356,.805
), nrow = 5, byrow = TRUE,
   dimnames = list(stages,stages)
)
print(A)


```

##Part a
To calculate $\lambda$ from the popbio package we do the following:
```{r}
library(popbio)
lambda(A)
```
Thus $\lambda = 1.04$.

##Part b
To calculate the stable stage distribution we do the following:
```{r}
stable.stage(A)
```

