---
title: "Population & Community Ecology Homework 1"
author: Samantha Apgar & Henry Frye
output: 
  pdf_document: default
---

#Problem 1

In this problem we look to find the density of duckweed plants per $m^2$ after 47 days. We will estimate the future population using the exponential growth formula, $N_t = N_0 e^{rt}$ where the initial population, $N_0$, is 11.1 duckweed plants per $m^2$.  We can calculate the rate of unitless population change, r, by using the above equation.

```{r}
r <- (log(23.5/11.1))/20
r
```

Thus with r= 0.03750, we can calculate $N_{47}$ by using the equation $N_t = N_0 e^{rt}$ again.

```{r}
N47 <- 11.1*(exp(0.0375*47))
N47
```

Thus after a total of 47 days, the population density will be 64.680 duckweed plants per $m^2$.

***

#Problem 2

In this problem we want to find the time at which both Ethipia and Mexico will have the same population size, while starting at different population sizes and growing at different rates. We will determine this time by setting the exponential growth formula, $N_t = N_0 e^{rt}$ for each population equal to eachother and solving for t (time). The initial population, $N_0$, for Mexico is 121,005,000 and r is 1.2% and the initial population for Ethiopia, $N_0$, is 90,076,000 and r is 2.9%. 

```{r}
121005000exp
```

In 0.1736 years the populations of Mexico and Ethiopia will be equal. 

***

#Problem 3

##Part a 
Given the information we would like to solve for $N_7$, the population in seven years.

```{r}
n_7 <- (1.037^7)*112
n_7
```

Thus the population density will be 144 (rounded, no units given) at year 7.

##Part b
Using the exponential growth formula, $N_t = N_0 e^{rt}$, we would like to
solve for an r value where $N_0 = 112$ and $N_t = 144$. Thus we solve for r as $ln(\frac{N_t}{N_0})/t=r$.  

```{r}
r <- log(144/112)/7
r
```

Thus r is equal to .036 in the exponential model to  achieve the same results as the discrete population growth model.

***

#Problem 4

***



#Problem 5

Survivorship, as defined by $\frac{S_x}{S_0}$, is calculated in the vector below.
```{r}
age <- 0:4
sx <- c(740,280,105,32,0)
bx <- c(0,0.4,1.3,3,8,NA)
lx <- sx/740
print(lx)
```

**I assume that g(x) was meant to be p(x)**

Solving for p(x)


***

#Problem 6


***

#Problem 7

The following is the setup for the population project matrix:
```{r}
stages <- c('seedling','juvenile','small adult', 'medium adult', 'large adult')
A <-  matrix(c(
     0,  0,.015,.063,.189,
  .294,.25, 0  , 0  , 0  ,
  .176,.35,.385,.059,.012,
  0   ,.35,.461,.568,.171,
  0   , 0 ,.108,.356,.805
), nrow = 5, byrow = TRUE,
   dimnames = list(stages,stages)
)
print(A)


```

##Part a
To calculate $\lambda$ from the popbio package we do the following:
```{r}
library(popbio)
lambda(A)
```
Thus $\lambda = 1.04$.

##Part b
To calculate the stable stage distribution we do the following:
```{r}
stable.stage(A)
```

