---
title: "Population & Community Ecology Homework 1"
author: Samantha Apgar & Henry Frye
output: 
  pdf_document: default
---

#Problem 1

In this problem we look to find the density of duckweed plants per $m^2$ after 47 days. We will estimate the future population using the exponential growth formula, $N_t = N_0 e^{rt}$ where the initial population, $N_0$, is 11.1 duckweed plants per $m^2$.  We can calculate the rate of unitless population change, r, by using the above equation.

```{r}
r <- (log(23.5/11.1))/20
r
```

Thus with r= 0.03750, we can calculate $N_{47}$ by using the equation $N_t = N_0 e^{rt}$ again.

```{r}
N47 <- 11.1*(exp(0.0375*47))
N47
```

Thus after a total of 47 days, the population density will be 64.680 duckweed plants per $m^2$.

***

#Problem 2

In this problem we want to find the time at which both Ethipia and Mexico will have the same population size, while starting at different population sizes and growing at different rates. We will determine this time by setting the exponential growth formula, $N_t = N_0 e^{rt}$ for each population equal to eachother and solving for t (time). The initial population, $N_0$, for Mexico is 121,005,000 and r is 1.2% and the initial population for Ethiopia, $N_0$, is 90,076,000 and r is 2.9%. 

```{r}
#SO. basically this problem has the same variable on both sides of the equation and I'm not sure how to turn that into a single line of code. If you can do that I may think you're a genius.

12100500*exp(1.2*t)=90076000*exp(2.9t)
log(exp(1.2t))=log(0.74440*exp(2.9t))
1.2t= -.29518 + 2.9t
t <- -0.29518/-1.7
t
```

In 0.1736 years the populations of Mexico and Ethiopia will be equal. 

***

#Problem 3

##Part a 
Given the information we would like to solve for $N_7$, the population in seven years.

```{r}
n_7 <- (1.037^7)*112
n_7
```

Thus the population density will be 144 (rounded, no units given) at year 7.

##Part b
Using the exponential growth formula, $N_t = N_0 e^{rt}$, we would like to
solve for an r value where $N_0 = 112$ and $N_t = 144$. Thus we solve for r as $ln(\frac{N_t}{N_0})/t=r$.  

```{r}
r <- log(144/112)/7
r
```

Thus r is equal to .036 in the exponential model to  achieve the same results as the discrete population growth model.

***

#Problem 4

***
In this problem, we are given values for the number of collared doves monitored in Great Britain each year. We are asked to estimate r by plotting the natural log of the number of doves by the time in years. 

```{r}
Time <- c(1955, 1956, 1957, 1958, 1959, 1960, 1961, 1962, 1963, 1964)
Time 
NumDoves <- c(4, 16, 45, 100, 205, 675, 1900, 4650, 10200, 18885)
NumDoves
LnNumDoves <- log(NumDoves)
LnNumDoves

lm(LnNumDoves~Time)

```
r = 0.9367

In the second part of this question, we are asked to plot ln(N) versus t:
```{r}
plot(Time, LnNumDoves)
```

We are also asked to calculate what the population sizes would be each year, assuming exponential growth, given that the initial population size in 1955 is 4 doves.

```{r}
TimeSinceStart <- c(1,2,3,4,5,6,7,8,9)
TimeSinceStart
ExponEqu <- 4*exp(0.9367*TimeSinceStart)
ExponEqu
for(val in Time)
print(val)
```

The population sizes would be as follows:
1955: 4 doves
1956: 10.20619 doves
1957: 26.04158 doves
1958: 66.44632 doves
1959: 169.54083 doves
1960: 432.59171 doves
1961: 1103.77827 doves
1962: 2816.34257 doves
1963: 7186.03157 doves
1964: 18335.50020 doves

#Problem 5

$l_x$ and $p_x$ are solved below and placed into the following life history table:
```{r}
library(knitr)
age <- 0:4
sx <- c(740,280,105,32,0)
bx <- c(0,0.4,1.3,3.8,NA)
lx <- sx/740

px <- vector(mode = 'numeric',length = 5)
for(i in 1:length(lx)){
  px[i] <- (lx[i+1]/lx[i])
}

lifetable <- cbind(age,sx,lx,px,bx)
colnames(lifetable) <- c('Age','S_x','l_x','p_x','b_x')
lifetable <- as.data.frame(lifetable)
kable(lifetable)


```

We solve for $R_0$ in the following:
```{r}
r0 <- sum(lifetable$l_x*lifetable$b_x, na.rm= TRUE)
r0
```

Thus $R_0= .5$

We solve for G as
```{r}
(with(lifetable, sum(age*l_x*b_x, na.rm = TRUE))/r0)

```

Thus we find the generation time, G, to be about 2.03.


***

#Problem 6

In this problem we are asked to use the life table from problem 5 to create a leslie matrix and project the number in each age class for the next two years. First, we provide the completed life table below:

```{r}
library(knitr)
age <- 0:4
sx <- c(740,280,105,32,0)
bx <- c(0,0.4,1.3,3.8,NA)
lx <- sx/740
lxbx <- lx*bx
agelxbx <- age*lx*bx

px <- vector(mode = 'numeric',length = 5)
for(i in 1:length(lx)){
  px[i] <- (lx[i+1]/lx[i])
}

pxbx1 <- vector(mode = 'numeric',length = 5)
for(i in 1:length(px)){
  pxbx1[i] <- (px*bx[i+1])
}

pxbx1

lifetable2 <- cbind(age,sx,lx,px,bx,lxbx,agelxbx,pxbx1)
colnames(lifetable2) <- c('Age','S_x','l_x','p_x','b_x', 'lx_bx', 'x_lx_bx', 'px_bx+1')
lifetable2 <- as.data.frame(lifetable2)
kable(lifetable2)

```

Next, we will construct the leslie matrix using $m_0$, $m_1$, $m_2$, $m_3$, $p_0$, $p_1$, and $p_2$. Then we will multiply the leslie matrix by the projection matrix two separate times to receive the projected number of individuals at each age class over the next two years. 

```{r}
leslie <- matrix(c(0.15136, 0.4875, 1.15824, 0, 0.3784, 0, 0, 0, 0, 0.3750, 0, 0, 0, 0, 0.3048, 0), nrow = 4, byrow = T)
leslie
leslie <- as.data.frame(leslie)
kable(leslie)
leslie

projection1 <- matrix(c(70, 40, 20, 10), nrow = 4, byrow = T)
projection1

# HELP ME MULTIPLY MATRICES IN R UGH- basically the leslie matrix needs to be multiplied by the projection1 matrix and then those results should be the projection2 matrix that is again multiplied by the leslie matrix

projection1output <- leslie %*% projection1

```


***

#Problem 7

The following is the setup for the population project matrix:
```{r}

stages <- c('seedling','juvenile','small adult', 'medium adult', 'large adult')
A <-  matrix(c(
     0,  0,.015,.063,.189,
  .294,.25, 0  , 0  , 0  ,
  .176,.35,.385,.059,.012,
     0,.35,.461,.568,.171,
     0, 0 ,.108,.356,.805
), nrow = 5, byrow = TRUE,
   dimnames = list(stages,stages)
)
```

##Population Projection matrix
```{r}
library(knitr)
kable(A)
```


##Part a
To calculate $\lambda$ from the popbio package we do the following:
```{r}
library(popbio)
lambda(A)
```
Thus $\lambda = 1.04$.

##Part b
To calculate the stable stage distribution we do the following:
```{r}
stable.stage(A)
```
##Part c
Calculating sensitivity table as such:
```{r}
sense <- sensitivity(A)
kable(sense)
```

Calculating elasticity table
```{r}
elast <- elasticity(A)
kable(elast)
```


The largest elasticity value is the probability of staying a large adult tree, with a value of .42.


##Part d
Cutting the value of adult tree fecundities by half:
```{r}
B <- A
.189/2 -> B[1,5]
.063/2 -> B[1,4]
.015/2 -> B[1,3]
lambda(B)
```

Thus the value of $\lambda$ is now 1.013 and the stable-age distribution is:

```{r}
stable.stage(B)
```


When the fecunditities for adult trees were cut by half lambda decreased by .024.  The stable-age distribution was higher for medium and adult trees when fecundities of all adult trees was reduced suggesting a forest with stable ages skewed towards older trees.

##Part e
Cutting the probabilities that trees stay in the same class by half:
```{r}
C <- A
.385/2 -> C[3,3]
.568/2 -> C[4,4]
.805/2 -> C[5,5]
lambda(C)
```

Thus the value of lambda is reduced quite a bit more than the previous case with a value of .746. The stable-age distribution is:
```{r}
stable.stage(C)
```

Now the stable age distribution has now increased in all classes but large adult trees. This suggests that large adult trees have a large role in dictating the age distribution within the population.

#Problem 8

```{r, echo = FALSE}

StochasticSim <- function(steps,b,d,N0){
  N <- numeric(steps + 1);
  t <- numeric(steps + 1);
  N[1] <- N0;
  t[1] <- 0;
  for(i in 1:steps){
    if (N[i]>0) {
      eventTime <- rexp(1,(b+d)*N[i])
      t[i+1] <- t[i] + eventTime 
    if (runif(1) < (b/(b+d))) {
      N[i+1] <- N[i] + 1
  } else {
    N[i+1] <- N[i] - 1
    }
  }
  else {
      N[i+1] <- 0
      t[i+1] <- t[i]
    }
}
#plot(N~t,type='l')
return(N[steps+1])
}
```

Running the stochastic simulation 50 times with 50 steps, $N_0=10$, b=.55 and d=.5 yields the following vector.
```{r}
sim10runs <- vector(mode = 'numeric',length = 50)

for(i in 1:50){
    sim10runs[i] <- StochasticSim(steps = 50,b=.55,d=.5, N0=10)
} 


propextinction10 <- length(which(sim10runs == 0))/length(sim10runs)
paste('Extinction rate is:',propextinction10)
sim10runs
```

Running the same stochastic simulation, but with $N_0=50$ yields

```{r}
sim50runs <- vector(mode = 'numeric',length = 50)

for(i in 1:50){
    sim50runs[i] <- StochasticSim(steps = 50,b=.55,d=.5, N0=50)
} 


propextinction50 <- length(which(sim50runs == 0))/length(sim50runs)
paste('Extinction rate is:',propextinction50)
sim50runs
```



#Problem 9

In problem number 9 we are asked to calculate the average weight of trees in Plot B and the total weight of plants in both Plots A and B. We are given the following values: Plot A: 16 seedlings/$m^2$, 1 tree = 24kg, 1 tree/$m^2$, Plot B: 4 seedlings/$m^2$ and 0.5 trees/$m^2$. We will use the equation w = $cN^k$ to first calculate c using the values from Plot A and then use the value of c to calculate the value of w in Plot B. 

```{r}
c <- 24/1^(-3/2)
c

w <- 24*(0.5^(-3/2))
w
```

The average tree in Plot B weighs 67.88225 kg 

The total weight of plants in each plot is 24 kg. 