---
title: "EEB 5301 Homework 4"
author: "Samantha Apgar & Henry Frye"
date: "Spring 2017"
output: 
  tufte::tufte_handout:
    keep_tex: false
---

# Problem 1

# Problem 2

```{r vegan setup, echo = FALSE, warning = FALSE, message = FALSE}
library(vegan)
library(dplyr)
library(tidyr)
data(dune)
data(dune.env)

dune.sum	<- aggregate(dune,	list(Moisture	=	dune.env$Moisture),	sum)
dune.sum1	<- dune.sum[,-1]
rownames(dune.sum1)	<- dune.sum[,1]

radPlot	<- function(x,	add=FALSE,	col='black',	...){
		x	<- x[x>0]
		x	<- x[rev(order(x))]
		if(!add)
				plot(1:length(x),	x,	type='l',	xlab='Rank',	ylab='Abundance	(log	scale)',
									log='y',	col=col,	...)
		else
				lines(1:length(x),	x,	col=col,	...)
		return(NULL)}

```

##Part a

```{r}
specnumber(dune.sum1)
```

As moisture increases, species richness decreases from moisture level 1 to level 4, but then suddenly increases at level 5. The most rich is level 5 while level 4 is the least rich.

## Part b

```{r}
sort(diversity(dune.sum1))
```

According to Shannon's Index, the moisture level is ranked 5 > 2 > 1 > 4 in order of high diversity to lowest.

## Part c

```{r}
sort(exp(diversity(dune.sum1)))
sort(1/(diversity(dune.sum1, index = 'invsimpson')))
```

The effective number of species between Simpson's and Shannon's index are almost flipped.  Moisture level 4 has the lowest effective number of species according to Shannon's index while the level 4 is considered the having the highest effective number of species according to Simpson's.

## Part d









