---
title: "R Warmup"
output: pdf_document
author: Henry Frye
---

#Objects and Arithmetic

##Exercise 1

###Part a
This will result in the length of vector x
```{r}
x<-c(4,2,6)
y<-c(1,0,-1)
length(x)

```

###Part b
This will result in the sum of vector x.
```{r}
sum(x)
```

###Part c
This will result in the sum of the square for each value within vector x.
```{r}
sum(x^2)
```

###Part d
This will result in the sum of each value between vectors x and y based on position within the vectors
```{r}
x+y
```

###Part e
This will result in the product of each value between vectors x and y based on position within the vectors
```{r}
x*y
```

###Part f
This will subtract 2 from each value in vector x.
```{r}
x-2
```

###Part g
This will square each value in vector x.
```{r}
x^2
```

##Exercise 2
###Part a
This will give a sequence of whole integers [7,11].
```{r}
7:11
```
###Part b
This will give a sequence of whole integers [2,9].
```{r}
seq(2,9)
```
###Part c
This will give a sequence of integers counting by 2 from 4 to 10.
```{r}
seq(4,10,by=2)
```
###Part d
This will give a sequence of integers between and 3 and 30 divided into 10 parts.
```{r}
seq(3,30,length=10)
```
###Part e
This will give a descending sequence from 6 to -4 counting from 2's.
```{r}
seq(6,-4,by=-2)
```

***
#Summaries and Subscripting
##Exercise 1
###Part a
This will return the second value of vector x.
```{r}
x<- c(5,9,2,3,4,6,7,0,8,12,2,9)
x[2]
```
###Part b
This will return the second through fourth values of vector x
```{r}
x[2:4]
```
###Part c
This will return the second, the third, and the sixth value of vector x.
```{r}
x[c(2,3,6)]
```

***
#Matrices
##Exercise 1
###Setup
```{r}
x <- matrix(c(3,2,-1,1), nrow = 2)
print(x)

y<- matrix(c(1,4,0,0,1,-1), nrow = 2)
print(y)

```

###Part a
This will multiply each value in matrix **x** by 2. 
```{r}
2*x
```

###Part b
This will result in the square of **x**. 
```{r}
x*x
```


###Part e
This will trnanspose matrix **y**.
```{r}
t(y)
```

##Exercise 2
###Part a

```{r}
x[1,]
```

Part b
```{r}
x[2,]
```

Part c
```{r}
x[,2]
```

Part d
```{r}
y[1,2]
```

Part e
```{r}
y[,2:3]
```

***
#Statistical Computation and Simulation
##Exercise 1
Part a
```{r}
dnorm(x = .5, mean = 2, sd = .25)
```

Part b
```{r}
pnorm(2.5, mean = 2, sd = .25)
```

Part c
```{r}
qnorm(.95, mean =2, sd = .25)
```

Part d
```{r}
pnorm(3, mean = 2, sd = .25) - pnorm(1, mean = 2, sd = .25)
```


##Exercise 3
```{r}
pois.sim <- rpois(n = 100, lambda = 5)
summary(pois.sim)
```


# Graphics
##Exercise 1
```{r}
gam.sim <- rgamma(n = 100, shape = .5)

par(mfrow = c(1,2))
hist(gam.sim, xlab = 'Simulated Gamma Values', col = 'orange',
     main = 'Gamma Histogram')
boxplot(gam.sim, notch = TRUE, col = 'cadetblue', main = 'Gamma Boxplot', main.cex = .5)

```


##Exercise 3

```{r}
data(nhtemp)
plot(nhtemp)
```

